---
title: "covid 19"
author: "Syi-yonnie Richard"
date: "10/06/21"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')  
kRdata$state<-as.factor(kRdata$state)  
kRdata$result<-as.factor(kRdata$result)  
kRdata$sex<-as.factor(kRdata$sex)  
kRdata$Sex<-as.factor(kRdata$Sex)  

```


### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well 

  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this
```{r}
barchartGC(~result + Sex,data=kRdata)
```



While observing the chart I can see their is a big difference of the dead and alive bar, which, there were more chances of survival than death and that more men have died from covid than women.  


```{r}
barchartGC(~result + Sex,data=kRdata, type="percent")

```


 Observing the second chart we can see women had a more successful survival than men while the percentage of a women surviving is 55-56% and about 40% for men. Overall women had a more success with surviving covid while men had an unsuccessful chance.  


`
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

 This result table shows us the correct percentages where at first we seen only the alive Female group was 56% and M was 44%. In the group titled "dead" F ranged at 36% and M ranged at 64%. A pretty big difference was shown. About 1.4 females were later deceased, while, only 3% males were deceased. Males in both charts and percentages ended up deceased double times the women.


## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The chi sq test shows and provides almost the same amount between men and women's percentages. Each result (Null) and (Observed Counts) provides us with being merely close by comparing in the first results with Females being 1831 (Observant Counts) and 1817 (Expected by Null) shows a 14 % drop but still being in the 1800 range, then with Males it shows a close result between the null and observed counts with the null results resulting in 1411 and 1425 showing a close result.  
()





```{r}
fisher.test(table1)
```

 The Fisher table shows a ration table between the odds of dying for males being (44/1411) and for females being (25/1831). The overall answer is divided coming into the answer of 2.3 being in the likely range of having the higher chance of dying. The confidence intervals are 1.4 to 3.9.




## Conclusions

As reading into to all of the data we can see in Korea,Korean women had a higher chance in surviving covid than men as showing low percentages and ratios of men having a rating as high as 64 and 44. Men were represented as low survivors throughout every piece of data listed.
